Exploring probability distributions for bivariate temporal granularities

### 2020-01-27 main.Rmd
smart meter section update - aligning it with how paper flows

<!-- # how data looks -->
Smart meters provide large quantities of measurements on energy usage for households across Australia. One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption in `general_supply_kwh` in every half hour from February 2012 to March 2014. \ref{fig:smart-meter} shows the data structure of the available data. It would be interesting to explore the energy consumption distribution for these customers and gain insights on their energy behavior which are lost either due to aggregation or looking only at coarser temporal units. The idea here is to show how looking at the energy consumption across different cyclic granularities together can help identify different behavioral patterns and identify the extreme and regular households. We take a random sample of 50 households to see how we can achieve that.

Data structure for visualization:

The variable `reading_datetime` is the index, which represents the time variable. The variable `customer_id` denoting different households represent the key of the tsibble and `general_supply_kwh` is the time series variable which needs to be analyzed. Data sets of the form <C1, C2, `general_supply_kwh`> will form the basis of exploration for each key in the data set, the cyclic granularity C1 mapped to x axis and `general_supply_kwh` to y-axis with the cyclic granularity C2 as the facet. Any geometry displaying the probability distribution of `general_supply_kwh` like  boxplot, letter value, violin,
ridge or highest density region plots can then be used to explore the data set.

Cyclic granularities search and computation:

While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.

@@

@@

**gravitas** R package [@R-gravitas] is used to facilitate the systematic exploration here.

The variable 'reading_datetime` is the index, which represents the time variable. While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.

These cyclic granularities are a function of the index variable and are computed based on if they are cicular (@ref:Sec), pseudo-circular (@ref:Sec) or aperiodic (@ref:Sec). If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, fortnight, month, quarter, semester, year), the following cyclic time granularities can be considered for this analysis. The interval of this tsibble is 30 minutes, and hence the temporal granularities may range from half-hour to year.

```{r}
library(gravitas)
smart_meter %>% search_gran()
```

@@

```{r}
smart_meter10 %>% 
  search_gran(highest_unit = "month")
```

@@

```{r search_gran_limit2}

@@

Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using \autoref{}.

@@

Figure \ref{fig:wknd-wday50} shows the quantile plot of energy consumption of 50 households. Similar quantiles for weekend and weekdays indicate that behaviors of most of these customers do not alter between weekdays and weekends or looking at all customers together balances out the fact

Figure \ref{fig:my-hd} shows that the median is very close to the lower boundaries of all the other bands implying energy consumption for these households are left skewed. Moreover, only the pink band changes significantly in winter months (May - August). The median consumption or width of other bands doesn't vary much across seasons, implying extreme behavior or extreme customers does not vary across seasons much. Most of the behavioral changes occur in the quartiles, that too in peak hours of the day. It is to be noted here that the the level of quantiles increased too in winter months (obvious because of weather conditions), the interesting part is to notice that the relationship between bands other than quartile stayed same across seasons.

@@

In Figure \ref{fig:day-fortnight}, the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy behavior of these two customers. it can be seen that for most days of the fortnight, the second household has much less consumption than the first one. However, there is additional information that we can derive looking at the distribution. If we consider letter value F as a regular behavior and letter values beyond F as not-so-regular behavior, we can conclude that the regular behavior of the first household is more stable than the second household. However, the distribution of tail of the first household is more variable, observed through distinct letter values, implying that their not-so-regular behavior is quite extreme. This shows, how looking at the distribution of the dependent variable can throw more light on the energy behavior of the customers, which are lost using aggregate or summary statistics.

>>>>

<!-- # how data looks -->
Smart meters provide large quantities of measurements on energy usage for households across Australia. One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption in `general_supply_kwh` in every half hour from February 2012 to March 2014. \ref{fig:smart-meter} shows the data structure of the available data. It would be interesting to explore the energy consumption distribution for these customers and gain insights on their energy behavior which are lost either due to aggregation or looking only at coarser temporal units. The idea here is to show how looking at the energy consumption across different cyclic granularities together can help identify different behavioral patterns and identify the extreme and regular households.

<!-- We take a random sample of 50 households to see how we can achieve that.  -->

### Data structure for visualization:

The variable `reading_datetime` is the index of the tsibble and represents the time variable. The variable `customer_id` denoting different households represent the key of the tsibble and `general_supply_kwh` is the time series variable which needs to be analyzed. Data sets of the form <C1, C2, `general_supply_kwh`> will form the basis of exploration for each key in the data set, where the cyclic granularity C1 would be mapped to x axis and `general_supply_kwh` to y-axis with the cyclic granularity C2 mapped to facet. Any geometry displaying the probability distribution of `general_supply_kwh` like  boxplot, letter value, violin,
ridge or highest density region plots can then be used to explore the data set.

@@

### Cyclic granularities search and computation:

**gravitas** R package [@R-gravitas] is used to facilitate the systematic exploration here. While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.

@@

These cyclic granularities are a function of the index variable and are computed based on if they are cicular (Section \ref{sec:circular-gran-def}), pseudo-circular (Section \ref{sec:quasi-circular-gran-def}) or aperiodic (Section \ref{sec:aperiodic-gran-def}). If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, fortnight, month, quarter, semester, year), the following cyclic time granularities can be considered for this analysis. The interval of this tsibble is 30 minutes, and hence the temporal granularities may range from half-hour to year.

```{r, echo=TRUE}
library(gravitas)
smart_meter %>% search_gran()
```

@@

```{r, echo=TRUE}
smart_meter10 %>% 
  search_gran(highest_unit = "month")
```

@@

```{r search_gran_limit2, echo=TRUE}

@@

### Screening and visualising harmonies:

Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using Sections \ref{sec:circular-gran-def}, \ref{sec:quasi-circular-gran-def} and \ref{sec:asperiodic-gran-def}.

@@

Each of these 13 pairs need to be visualised to have multiple perspectives on the data. If the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one `customer_id` shows a particular behavior, but this behavior is reversed when all keys are combined together. This is also intuitive because `customer_id`'s with very different occupation or demographics will tend to have very different energy behavior and combining them together will somehow weaken any typical or extreme behavior. For example, Figure \ref{fig:wknd-wday50} shows the quantile plot of energy consumption of a sample of 50 households. Quantiles for weekend and weekdays for these customers look exactly the same. It is not clear if the behaviors of most of these customers do not alter between weekdays and weekends or looking at all customers together balances out the behaviors. Consider another example in Figure \ref{fig:my-hd} showing the area quantile plots of 50 customers. The median is very close to the lower boundaries of all the other bands implying energy consumption for these households are left skewed. The pink band changes significantly in winter months (May - August) implying that much of the behavioral changes occur in the quartiles, that too in peak hours of the day. It is to be noted here that the the level of all quantiles increased too in winter months (increased energy usage in weather conditions), the interesting part is to notice that the relationship between bands other than quartile stayed same across seasons.

<!-- The median consumption or width of other bands doesn't vary much across seasons, implying extreme behavior or extreme customers does not vary across seasons much.  -->

@@

Since we established that looking at random samples with different demographics might confound our analysis, next the smart meter data is filtered for two customers in Figure \ref{fig:day-fortnight},to illustrate what kinds of insights can be drawn for the energy behavior of these two customers. it can be seen that for most days of the fortnight, the second household has much less consumption than the first one. However, there is additional information that we can derive looking at the distribution. If we consider letter value F as a regular behavior and letter values beyond F as not-so-regular behavior, we can conclude that the regular behavior of the first household is more stable than the second household. However, the distribution of tail of the first household is more variable, observed through distinct letter values, implying that their not-so-regular behavior is quite extreme. This shows, how looking at the distribution of the dependent variable can throw more light on the energy behavior of the customers, which are lost using aggregate or summary statistics.
